### 2018-11-06 Chapters/3.Hypothesis/Hypothesis.tex

\chapter{A Dynamic Programming Approach}

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=\textwidth]{Figures/ribbon-3d.pdf}
%     \caption[3D Ribbon Image]{An illustration of the parameters for a ribbon; the $r$ axis captures the thickness of the ribbon and the $s$ axis captures the distance along the initial ribbon. The $t$ axis records horizontal displacements from the initial ribbon's medial curve.}
%     \label{fig:ribbon_3d}
% \end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{Figures/ribbon-3d-combined.pdf}
    \caption[3D Ribbon Image]{An illustration of the parameters for a ribbon: (top) the $k$ axis captures the thickness of the ribbon and the $i$ axis captures the distance along the initial ribbon, the $j$ axis records horizontal displacements from the initial ribbon's medial curve; (bottom) the boundaries of the ribbon must lie within $\pm \MaxDistance{}$, pixels within $\pm \MinRadius{}$ of the center are probably part of the ribbon and pixels further than $\pm \MaxRadius{}$ are likely negative. }
    \label{fig:ribbon_3d}
\end{figure}

\section{The Ribbon Image}

In order to simplify the problem, we first create a straightened ``ribbon" image in which the
sidewalk runs in a straight line, even if the original sidewalk is curved. We represent a ribbon of
length $L$ using parameters along and across to a medial curve $\mathbf{x}(s)$, where $s \in [0,
L]$. The ribbon image is the set of all points within a perpendicular distance $r(s)$ to the medial
curve.

A second parameter, which we call $t$, is measured in a direction orthogonal to the curve. We define
ribbon coordinates $\mathbf{x}(s, t) = \mathbf{x}(s) + t\mathbf{n}(s)$, where $\mathbf{n}(s)$ is a
vector perpendicular to the medial curve at position $s$, which can be obtained by rotating the
curve's tangent vector $\dot{\mathbf{x}}(s)$ by 90 degrees.

In practice we represent the curve $\mathbf{x}(s)$ discretely as an arc-length re-sampled set of
points $\mathbf{x}_s,$ for $s=0,1,...,L-1$ from poly-line features that are read from a GIS file.
The points are sampled at an arc-length distance of about one pixel between each point, and then
Laplacian smoothing is applied in order diffuse the curvatures along the re-sampled poly-line. We
use a secant to estimate the tangent to the curve,
\[\dot{\mathbf{x}}_s = \left[\begin{array}{c} \dot{x}_s \\ \dot{y}_s \end{array}\right]= \frac{\mathbf{x}_{s+1}-\mathbf{x}_{s-1}}{||\mathbf{x}_{s+1}-\mathbf{x}_{s-1}||} \] and rotate it ninety degrees counter-clockwise in order to find a direction  $\mathbf{n}_s$, so that 
\[\mathbf{n}_s = \left[\begin{array}{c} -\dot{y}_s \\ \dot{x}_s \end{array}\right]. \]

\begin{figure}[H]
    \centering \includegraphics[width=\textwidth]{Figures/sw-figure.pdf} \caption[2D Ribbon Image]{
    A demonstration of the parameters for a ribbon in 2D. $k_n$ indicates a parameter that describes
    the radius of the medial curve that we considered to be part of the ribbon, $k_f$ is a distance
    from the medial curve that we considered very unlikely to be part of the ribbon, and $k_W$ is
    the maximum width of the ribbon image.} \label{fig:2d_ribbon}
\end{figure}

As shown in figure \ref{fig:2d_ribbon}, from figure \ref{fig:osm_oxford_path}, we should assume that
there is some uncertainty in the position $\mathbf{x}_s$ of the medial curve and also uncertainty in
the radius of the ribbon. But that there is range of distances $t \in 0\dots k_n$
from the medial curve which we can accept labeling as positive examples that are within the ribbon,
and that there is some range of distances $t \in k_f\dots k_W$ which we can accept labeling as
negative sample. Finally a range $t\in (k_n+1)...(k_f-1)$ of distances that are possible locations
of the ribbon's boundary. Therefore, when searching for improved medial curves and radii of the
ribbon, we can consider a discrete range of offsets $t=-k_W,..., k_W$ chosen to be wide enough that
true ribbon image itself likely fits entirely within the grid of input negative samples along its
edges.

The orthorectified color image $\InputImage(x,y)$ records the color at the point $x,y$ in the image
plane. Then $\InputImage(\mathbf{x}_{s,t}),$ is a warped version of the image in which the ribbon is
centered on a single strip at $t=0$. The situation is illustrated in figure  \ref{fig:ribbon_3d}. In
this warped image, we can estimate the amount \OutputTrajectory{} to shift the ribbon's center in
the $t-$axis, which is orthogonal to the curve,  and also the new radius \OutputRadius{} so
that the  ribbon at each row is in columns $(\OutputTrajectory{} -\OutputRadius{}
)\dots(\OutputTrajectory{}+\OutputRadius{})$. Our problem is to estimate the most likely values of
$\OutputTrajectory{}$ and $\OutputRadius{}$.

\section{Initial Density Estimation}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/gmm_sample_1.png}
    \caption[\ac{GMM} Result 1]{Example of GMMs on sample sidewalk. The first row and third row shows the foreground GMM for sample sidewalk, and the second row and fourth row shows the back-ground for simple sidewalk. We use grey-scale to show the changes of probabilities, where for foreground, white is more likely to be sidewalk, and black is not.}
    \label{fig:GMM_result}
\end{figure}

We aim to find a segmentation that combines \textit{shape} priors on the curvature and thickness of the ribbon-like features and also maximizes the likelihood of observing colors in the image. 
A variety of approaches exist to learn the conditional probability $\Pr(\Color{}|R)$ of observing a color ($\Color{}$)
 conditioned on the event ($R$) that a pixel is within a ribbon \cite{hartigan1979algorithm, jordan1999introduction}.
Many of these methods take as input a set of training examples in order to choose parameters for some model.
Our conjecture is that the initial estimate of our sidewalk's mid-line, while perhaps not perfect, is much more likely to cover pixels that are part of a sidewalk than the pixels at the edge of the ribbon image. 
We use this assumption and Bayes theorem to estimate the probability that a pixel is part of a ribbon conditioned on its color.

Let $k_n$ be a parameter that describes the radius around the medial curve that we will consider to be part of the ribbon with very high confidence. 
Let $k_f$ indicate a distance from the medial curve that we consider very unlikely to be part of the ribbon. 
Finally, let $k_W$ indicate maximum value of $t$ to consider when learning the distribution of colors in a ribbon image. 
We model the distribution of colors within the sidewalk as a Gaussian Mixture Model of $k_g=4$ full-co-variance gaussian distribution and we model the distribution of colors outside of the ribbon as a mixture $2k_g$ Gaussian since we expect more variation and complexity in this zone. 
Using the \ac{EM-GMM} algorithm, we are able to estimate $p(\Color{}|R)$ and $p(\Color{}|\neg R)$. 
Then $p(R|\Color{})=\frac{p(\Color{}|R)p(R)}{p(\Color{}|R)+p(\Color{}|\neg R)}$, where $p(R)$ is constant. 
As shown in figure \ref{fig:GMM_result}, given an initial estimate for the set of colors inside or outside of the ribbon, we estimate the weights and co-variances of the Gaussians using the \ac{EM-GMM} \cite{sridharan2014gaussian} algorithm.

\section{A Lattice Representation of the Problem} 
We represent the solution to this problem as graph where each node represents a point that may lie on the middle of a ribbon with a different width, and edges connect each node to a node that may have preceded it on a path.

We will assume that widths may take on values from a set 
of predefined odd numbers (so that there is always a pixel in the center) 
Then the negative log-likelihood $\rownll(s, j, k)$ of for a 
particular offset $j$ and radius $k$ at a single row $(s)$ is

\begin{align}
\rownll(s, j, k) &= -\log \Pr(k, j)  + \sum_{t} \pixelnll(s,t, j, k) \\
\pixelnll(s, t, j, k) &= 
\begin{cases}
 -\ln\Pr(\InputImage(\x_{s,t})|R) &\text{if } t\in [j-k, j+k]\\
 -\ln\Pr(\InputImage(\x_{s,t})|\neg R) &\text{if } t\not\in[j-k, j+k]
\end{cases}
\end{align}

which combines the probabilities of seeing each color at offset $s$ in the ribbon image conditioned on whether the color is inside or outside of the ribbon.
However, we assume ribbons are continuous with respect to the position of their mid-line and their thickness, likelihood of the most probable path up to offset $s$ is

\begin{align}
\ribbonnll(s, j, k) &=  \begin{cases}
-\ln \Pr(k, j) & \text{if } s = 0\\
   \rownll(s, j, k){+} \\
   \hspace{1em}\displaystyle\max_{\delta_t,\delta_r} 
   \left[
   -\ln p(\delta_t{,}\delta_r){+}
   \ribbonnll(s{-}1{,}j{+}\delta_t{,}k{+}\delta_r)
 \right]
 & \text{if } s > 0.\\
\end{cases}
\end{align}

The quantity $-\ln p(\delta_t, \delta_r)$ is a penalty for changing direction or thickness, and could be set to the negative log likelihood of observing a change $\delta_t$ in position and a change $\delta_r$ in thickness.

% \begin{algorithm}
% \caption{A Dynamic programming approach predict edges for ribbon-like feature}

% \begin{algorithmic}
% \State Initialize a matrix $A$ to same shape as ribbon image
% \For {each line of pixel $s$ in ribbon image}
%     \For{each possible pixel as $j$ in width of line}
%         \For{each pixel as $k$ in between $j$ and edge of line}
%             \State Store the returned probabilities from function D in $A[0, ts, rs]$
%             \If{not the first line pixel}
%                 \For {each $\delta_t and \delta_r \in \{-1, 0, 1\}$}
%                     \State Store max sum value from returned probability,
%                     \State $A[s-1, j+\delta_t, k+\delta_r]$ and penalty in $A[s, j, k]$
%                     \State Also store the value of $\delta_t$ and $\delta_r$
%                 \EndFor
%             \EndIf
%         \EndFor
%     \EndFor
% \EndFor

% \end{algorithmic}
% \end{algorithm}